---
title: "p8105_hw3_yy3296"
author: "Yuze Yuan"
date: "2022-10-12"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```


#### Answer questions about the data

This dataset contains 1384617 rows and 15 columns, with each row
resprenting a single product from an instacart order. Variables include
identifiers for user, order, and product; the order in which each
product was added to the cart. There are several order-level variables,
describing the day and time of the order, and number of days since prior
order. Then there are several item-specific variables, describing the
product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs,
                                                 produce), and aisle (e.g. yogurt, fresh fruits), and whether the item
has been ordered by this user in the past. In total, there are 39123
products found in 131209 orders from 131209 distinct users.

Below is a table summarizing the number of items ordered from aisle. In
total, there are 134 aisles, with fresh vegetables and fresh fruits
holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```


Next is a plot that shows the number of items ordered in each aisle.
Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


Our next table shows the three most popular items in aisles
`baking ingredients`, `dog food care`, and `packaged vegetables fruits`,
and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

  
Finally is a table showing the mean hour of the day at which Pink Lady
Apples and Coffee Ice Cream are ordered on each day of the week. This
table has been formatted in an untidy manner for human readers. Pink
Lady Apples are generally purchased slightly earlier in the day than
Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

### Problem 2

I load the accel dataset and tidy the data, including transform Monday-Sunday to Weekday or Weekend, and encode data with reasonable variable classes. 

```{r}
clean_accel_data =
  read_csv("./dataset/accel_data (3).csv") %>%
  janitor::clean_names() %>% 
  mutate(
    weekday_weekend = ifelse((day == "Monday"| day == "Tuesday"|day == "Wednesday"|day == "Thursday"| day == "Friday"), "Weekday", "Weekend")) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_minutes",
    names_prefix = "activity_",
    values_to = "activity_counts") 
```
* Distribution
  * There are `r ncol(clean_accel_data)` variables in the dataset, which are `r colnames(clean_accel_data)`. 
  * There are `r nrow(clean_accel_data)` observations exists in the dataset. 
  
I aggregate across minutes to create a new variable clalled activity_whole_day to show the total activity for each day, and then I create a table showing these totals. 

```{r}
clean_accel_data %>%
  group_by(week, day) %>%
  summarize(activity_whole_day = sum(activity_counts)) %>%
  knitr::kable()
```

* I don't think there are any trends apparent.

I make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. 

```{r}
clean_accel_data %>%
  mutate(activity_minutes = as.integer(activity_minutes)) %>%
  ggplot(aes(x = activity_minutes, y = activity_counts, color = day)) + geom_point(alpha = 0.5) + labs(x = "activity minutes", y = "activity counts", title = "24-hour activity time courses for each day") + 
scale_x_continuous(
    breaks = seq(0, 1440, 120))
```

* Distribution
  * For this plot, we can see the largest data appears between 1080th and 1200th minutes, and the smalles data appears before about 300th mimutes. 
